---
title: "5206 Lecture-01"
author: "Ziying Liu"
date: "12/4/2019"
output: pdf_document
---

```{r setup, include=FALSE}

```
**Warm up**
\newline
Data structures
\newline
Homoheneous : Atomic vector; Matrix ; Array
Heterogeneous : List; Data Frame 
```{r}
x = c(5,29,13,87)
str(x) #Best way to understand what data structures is (str() (short for structure))

```

**Vector**
Two flavours: atomic vectors; lists
\newline
Three common properties:
\newline
1. Type, typeof(), what it is.
\newline
2. Length, length(), how many elements it contains.
\newline
3.  Attributes, attributes(), additional arbitrary metadata.


```{r}
NA^0
str(c("a", 1))
```
**Attributes**
\newline
Attributes can be retrieved/modified: individually with attr(), or “En masse” with attributes()/structure().

```{r}
a <- 1:3
attr(a, "x") <- "abcdef"
attr(a, "y") <- 4:6

attr(a,'x')
attributes(a)
str(attributes(a))

#or equivalently
b <- structure(
1:3,
x = "abcdef",
y = 4:6
)
b
```

```{r}
x <- 1:3
names(x) <- c("a", "b", "c")  #By assigning a character vector to names()

y <- setNames(1:3, c("a", "b", "c"))
#Remove names with unname(x) or names(x) <- NULL.
```

**Matrix**
```{r}
a <- matrix(1:6, nrow = 2, ncol = 3) # two dimensions matrix
a
rownames(a) = c('x','y')
b <- array(1:12, c(2, 3, 2)) #multi-dimensional array
b

```
Functions for working with vectors, matrices and arrays:
Vector        |Matrix        |Array
------------- | -------------|------------
names()       | rownames(), colnames()| dimnames()
length()      |nrow(), ncol()| dim()
c()           |rbind(),cbind()|abind::abind()
-             |t()           |aperm()
is.null(dim(x))| is.matrix() |is.array()



**Factor**
```{r}
x <- factor(c("a", "b", "b", "a"))
x
attributes(x)
table(x)

grade <- ordered(c("b", "b", "a", "c"), levels = c("c", "b", "a"))
grade #Ordered factors
```
**Dates**
```{r}
today <- Sys.Date()
attributes(today)
unclass(today) #the number of days since 1970-01-01

```

**Dates-times**
Two ways of storing this information: POSIXct, and POSIXlt.
\newline
Odd names:
\newline
1. “POSIX” is short for “Portable Operating System Interface”,
\newline
2. “ct” stands for calendar time (the time_t type in C),
\newline
3. and “lt” for local time (the struct tm type in C).
\newline
Focus on POSIXct (the simplest):
\newline
1. Built on top of a double vector.
\newline
2. Value = number of seconds since 1970-01-01.
```{r}
now_ct <- as.POSIXct("2018-08-01 22:00", tz = "UTC")
now_ct
attributes(now_ct)
structure(now_ct, tzone = "Asia/Tokyo")
one_week_1 <- as.difftime(1, units = "weeks")
one_week_1
attributes(one_week_1)

one_week_2 <- as.difftime(7, units = "days")
one_week_2

attributes(one_week_2)
```
**List**
```{r}
l1 <- list(
1:3,
"a",
c(TRUE, FALSE, TRUE),
c(2.3, 5.9)
)
l1
str(l1)
l4 <- list(list(1, 2), c(3, 4))
l5 <- c(list(1, 2), c(3, 4)) # c() will combine several lists into one:
str(l4)
str(l5)
```

**Data frame and tibbles**
```{r}
df1 <- data.frame(x = 2:4, y = letters[1:3])
attributes(df1)
df1

library(tibble) #tibble, a modern reimagining of the data frame.
df2 <- tibble(x = 4:6, y = letters[1:3])
typeof(df2)
attributes(df2)

```
List columns
```{r}
df <- data.frame(x = 1:3)
df$y <- list(1:2, 1:3, 1:4)
data.frame(
x = 1:3,
y = I(list(1:2, 1:3, 1:4))
)
tibble(
x = 1:3,
y = list(1:2, 1:3, 1:4)
)
```


```{r}
vals <- outer(1:5, 1:5, FUN = "paste", sep = ",")
vals
vals[c(4,8)]

select <- matrix(ncol = 2, byrow = TRUE, c(
1, 1,
3, 1,
2, 4
))
vals[select]


```



```{r}
df = tibble(x = 1:3, y = 3:1, z = letters[1:3])
df
str(df["x"])
str(df[, "x"])


```
```{r}
x <- list(1:3, "a", 4:6)
x
```
































